Designing Types for R, Empirically

### 2020-10-12 paper/main.tex
background

\subsection{Related Work}

@@

But what if the design of the type system is unclear?  \citet{tip} propose
an intriguing approach called trace typing. With trace typing, a new type
system can be prototyped and evaluated by applying the type rule to
execution traces of programs. While the approach has the limitation of
dynamic analysis techniques, namely that the results are only as good at the
coverage of the source code, it allows one to quickly test new design and
quantify how much of a code base can be type-checked.  Other approaches that
infer types for dynamic analysis include the work of \citet{FurrAF2009} and
\citet{hicks2011ruby} for Ruby.

@@

\subsection{The R Programming Language} %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

The R Project is a key tool for data analysis.  At the heart of the R
project is a vectorized, dynamic, lazy, functional, object-oriented
language with a rather unusual combination of features~\cite{ecoop12}
designed to ease learning by non-programmers and enable rapid
development of new statistical methods.  The language was
designed~\citet{R96} as a successor to S~\cite{S88}.

@@

\begin{lstlisting}
f <- function(x, ..., y=if(z==0) 1, z=0) { 
  x + y + if (missing(...)) 0 else c(...)
}
\end{lstlisting}

>>>>

\subsection{Related Work} %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

But what if the design of the type system is unclear?  \citet{tip} propose
a promising approach called trace typing. With trace typing, a new type
system can be prototyped and evaluated by applying the type rule to
execution traces of programs. While the approach has the limitation of
dynamic analysis techniques, namely that the results are only as good at the
coverage of the source code, it allows one to quickly test new design and
quantify how much of a code base can be type-checked.  Other approaches that
infer types for dynamic analysis include the work of \citet{FurrAF2009} and
\citet{hicks2011ruby} for Ruby.

@@

\subsection{The R Language} %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

The R Project is a key tool for data analysis.  At the heart of the R
project is a vectorized, dynamic, lazy, functional, object-oriented
language with an unusual combination of features~\cite{ecoop12}.  R
was designed~\citet{R96} as a successor to S~\cite{S88}.

@@

\begin{lstlisting}
f <- function(x,..., y=if(z==0) 1, z=0) { x + y + if(missing(...)) 0 else c(...) }
\end{lstlisting}
